Localization properties, observable character and computational reliability of the linear response in open systems

### 2018-06-25 AnalyticStructure_main.tex
update

The Liouvillian superoperator $\Liouv$ for self-consistent systems is expressed as the sum of the unperturbed part plus a coupling term. The left action of these 
terms reads, respectively
\be\lb{LiouZeroDef1}
\Liouvnot \op O = \commutator{\hnot}{O} \;,
\ee
and
\be\lb{CouplDef1}
\coupl\op O = \commutator{\op V'[\op O]}{\dmnot}  = \op Q_0\op V'[\op O]\dmnot - \dmnot\op V'[\op O]\op Q_0 \;,
\ee
where $\op V'$ expresses the perturbation induced by the density dependence on ground state Hamiltonian and can be conveniently expressed by defining the 
\emph{scalar coupling kernel}
\be
U\left[\op O; \op O'\right] \equiv  \int \dd \r \dd \r' \trace{\op O \frac{\delta \hnot }{\delta \rho_0(\r,\r')}
} \trace{\op O' \ket{\r} \bra{\r'}} \;, \nn
\ee
on the basis of which
\be\lb{VpDef1}
\op V'[\op O]= 
\int \dd \r \dd \r' U\left[\ketbra{\r}{\r'};\op O\right] \ketbra{\r'}{\r}\;.
\ee
In view of building the spectral decomposition we are interested in finding both the left and right eigenstates of $\Liouv$.   
The right action of $\Liouv$ can be assessed through the equivalence $\trace{\op O'(\Liouv\op O)} = \trace{(\op O'\Liouv)\op O}$, so that:
\be
\op O\Liouv = -\commutator{\hnot}{\op O} + \int \dd\r\dd\r'
U\left[\commutator{\dmnot}{\op O};\ketbra{\r'}{\r}\right]\ketbra{\r}{\r'} \;. \nn
\ee
In the cases relevant for the present analysis the coupling kernel is symmetric with respect to the exchange of its arguments so also the right action of the 
Liouvillian is expressed in terms of $\op V'$, as follows
\begin{align}\lb{LiouvillianRightActionDef1}
\op O\Liouv &= -\commutator{\hnot}{\op O} + V'\left[\commutator{{\dmnot}}{\op O}\right] = -\commutator{\hnot}{\op O} + \nn \\ 
&+ \dmnot\op V'\left[\commutator{{\dmnot}}{\op O}\right]\op Q_0 + \op Q_0\op V'\left[\commutator{\dmnot}{\op O}\right]\dmnot \;.
\end{align}
In the last equivalence we have made usage of the transverse property (see equation \eqref{RhopTransverseDef1}) of the coupling superoperator. Formula 
\eqref{LiouZeroDef1},\eqref{CouplDef1} and \eqref{LiouvillianRightActionDef1} are the starting point for the solution of the eigenvalues equations \eqref{ExcitationOperatorsDef1}.

@@

From the equations of motion for the excited states to the Casida equations...

>>>>

The Liouvillian superoperator $\Liouv$ for self-consistent systems is expressed as the sum of the unperturbed part plus a coupling term. The left action of these 
terms reads, respectively
\be\lb{LiouZeroDef1}
\Liouvnot \op O = \commutator{\hnot}{O} \;,
\ee
and
\be\lb{CouplDef1}
\coupl\op O = \commutator{\op V'[\op O]}{\dmnot}  = \op Q_0\op V'[\op O]\dmnot - \dmnot\op V'[\op O]\op Q_0 \;,
\ee
where $\op V'$ expresses the perturbation induced by the density dependence on ground state Hamiltonian and can be conveniently expressed by defining the 
\emph{scalar coupling kernel}
\be\lb{CouplingKernelDef1}
U\left[\op O; \op O'\right] \equiv  \int \dd \r \dd \r' \trace{\op O \frac{\delta \hnot }{\delta \rho_0(\r,\r')}
} \trace{\op O' \ket{\r} \bra{\r'}} \;, \nn
\ee
on the basis of which
\be\lb{VpDef1}
\op V'[\op O]= 
\int \dd \r \dd \r' U\left[\ketbra{\r}{\r'};\op O\right] \ketbra{\r'}{\r}\;.
\ee
In view of building the spectral decomposition we are interested in finding both the left and right eigenstates of $\Liouv$.   
The right action of $\Liouv$ can be assessed through the equivalence $\trace{\op O'(\Liouv\op O)} = \trace{(\op O'\Liouv)\op O}$, so that:
\be
\op O\Liouv = -\commutator{\hnot}{\op O} + \int \dd\r\dd\r'
U\left[\commutator{\dmnot}{\op O};\ketbra{\r'}{\r}\right]\ketbra{\r}{\r'} \;. \nn
\ee
In the cases relevant for the present analysis the coupling kernel is symmetric with respect to the exchange of its arguments so also the right action of the 
Liouvillian is expressed in terms of $\op V'$, as follows
\begin{align}\lb{LiouvillianRightActionDef1}
\op O\Liouv &= -\commutator{\hnot}{\op O} + V'\left[\commutator{{\dmnot}}{\op O}\right] = -\commutator{\hnot}{\op O} + \nn \\ 
&+ \dmnot\op V'\left[\commutator{{\dmnot}}{\op O}\right]\op Q_0 + \op Q_0\op V'\left[\commutator{\dmnot}{\op O}\right]\dmnot \;.
\end{align}
In the last equivalence we have made usage of the transverse property (see equation \eqref{RhopTransverseDef1}) of the coupling superoperator. Formula 
\eqref{LiouZeroDef1},\eqref{CouplDef1} and \eqref{LiouvillianRightActionDef1} are the starting point for the solution of the eigenvalues equations \eqref{ExcitationOperatorsDef1}.

@@

% We show that Casida equations are equivalent to the operator eigenvalue equations \eqref{ExcitationOperatorsDef1} when an explicit basis $\{\ket{\eta}\}$ for the empty states subspace is used,
% so that $\op Q_0 = \sum_{\eta}\ketbra{\eta}{\eta}$. Operators $\op O$ satisfying the transverse condition $\op O = \op O_{\perp}$ (see eq. \eqref{RhopTransverseDef1}) can be represented as 
% follows
% \be\lb{OperatorTransBasisExpDef1}
% \op O = \sum_{p\eta}\left(
% \excite{p}{\eta}\trace{\decay{\eta}{p}\op O} + \decay{\eta}{p}\trace{\excite{p}{\eta}\op O}
% \right) \;,
% \ee
% where the \emph{transition operators} $\excite{p}{\eta} \equiv \ketbra{\eta}{\psi_p}$ and $\decay{\eta}{p} \equiv \ketbra{\psi_p}{\eta}$
% are normalized as follows
% \be
% \trace{\excite{p}{\eta}\excite{q}{\rho}} = \trace{\decay{\eta}{p}\decay{\rho}{q}} = 0 \;, \quad
% \trace{\excite{p}{\eta}\decay{\rho}{q}} = \delta_{pq}\delta_{\eta\rho} \;. \nonumber
% \ee
% Expanding equations \eqref{ExcitationOperatorsDef1} in the basis of transitions provides a matrix equation for the coefficients of the expansion. In particular from the left action equation
% one has
% \be
% \sum_{q\rho}\mat{\trace{\decay{\alpha}{p}\, \sop{\mathcal L}\, \excite{q}{\beta}} &
% \trace{\decay{\alpha}{p}\, \sop{\mathcal L}\, \decay{\beta}{q}} \\
% \trace{\excite{p}{\alpha}\, \sop{\mathcal L}\, \excite{q}{\beta}} &
% \trace{\excite{p}{\alpha}\, \sop{\mathcal L}\, \decay{\beta}{q}}}
% \mat{X_a^{p\eta} \\ Y_a^{p\eta}} = \Omega_a \mat{X_a^{p\eta} \\ Y_a^{p\eta}}
% \ee

We show that Casida equations are equivalent to the equations of motion for the excited states \eqref{ExcitationOperatorsDef3}. To this scope we introduce an explicit basis $\{\ket{\eta}\}$
in the subspace of empty states so that both $\ket{\phi^a_p}$ and $\bra{\chi^a_p}$ can be represented as
\be
\ket{\phi^a_p} = \sum_\eta X^a_{p\eta}\ket{\eta} \;, \q
\bra{\chi^a_p} = \sum_\eta}\bra{\eta}Y^a_{p\eta} \;.
\ee
Plugging this expansion into equations \eqref{ExcitationOperatorsDef3} and projecting on a arbitrary element $\ket{\rho}$ of the basis provides a linear systems of equations for the coefficients
$X^a_{p\eta}$ and $Y^a_{p\eta}$, namely
\begin{align}
 \sum_\eta(H_{0\rho\eta}-\eps_p\delta_{\rho\eta})X^a_{p\eta} + \bra{\rho}\op V'[\op E_a]\ket{\psi_p} &= \Omega_a X^a_{p\rho} \;, \nn \\
 -\sum_\eta (H_{0\eta\rho}-\eps_p\delta_{\rho\eta})Y^a_{p\eta} - \bra{\psi_p}\op V'[\op E_a]\ket{\rho}& = \Omega_a Y^a_{p\rho} \;. \nn
\end{align}
Using this basis, excitation operators are expressed as a linear combination of transition operators $\excite{p}{\eta} \equiv \ketbra{\eta}{\psi_p}$ and $\decay{\eta}{p} \equiv \ketbra{\psi_p}{\eta}$,
as follows
\be\lb{ExcitationOpBasisTransition1}
\op E_a = \sum_{p\eta}\excite{p}{\eta}X^a_{p\eta}+\decay{\eta}{p}Y^a_{p\eta} \;.
\ee
The contribution of the coupling operator in the equations of motion of excited states can expressed by means of the coupling kernel \eqref{CouplingKernelDef1} as
\begin{align}
& \bra{\rho}\op V'[\op E_a]\ket{\psi_p} = \sum_{q\eta}\left(U[\decay{\rho}{p};\excite{q}{\eta}]X^a_{q\eta} + U[\decay{\rho}{p};\decay{\eta}{q}]Y^a_{q\eta}\right) \;,\nn \\
& \bra{\psi_p}\op V'[\op E_a]\ket{\rho} = \sum_{q\eta}\left(U[\excite{p}{\rho};\excite{q}{\eta}]X^a_{q\eta} + U[\excite{p}{\rho};\decay{\eta}{q}]Y^a_{q\eta}\right) \;, \nn
\end{align}
and so equations \eqref{ExcitationOperatorsDef3} can be recasted in matrix form
\be\lb{ExcitationMatrixEq1}
\sum_{q\eta}\mat{F_{p\rho}^{q\eta} &  D_{p\rho}^{q\eta}  \\
- E_{p\rho}^{q\eta} & - G_{p\rho}^{q\eta} }
\mat{X^a_{q\eta} \\ Y^a_{q\eta}} = \Omega_a \mat{X^a_{p\rho} \\ Y^a_{p\rho}}
\ee
where
\begin{align}
& F_{p\rho}^{q\eta} = (H_{0\rho\eta}-\eps_p\delta_{\rho\eta})\delta_{pq} + U[\decay{\rho}{p};\excite{q}{\eta}] \;,\nn \\
& D_{p\rho}^{q\eta} = U[\decay{\rho}{p};\decay{\eta}{q}] \;,\nn \\
& E_{p\rho}^{q\eta} = U[\excite{p}{\rho};\excite{q}{\eta}] \;,\nn \\
& G_{p\rho}^{q\eta} = (H_{0\eta\rho}-\eps_p\delta_{\rho\eta})\delta_{pq} + U[\excite{p}{\rho};\decay{\eta}{q}] \;. \nn
\end{align}
Finally, exploiting the symmetry properties of both the unperturbed Hamiltonian and of the coupling kernel it follows that $F=G$ and $D=E$ so that equations
\eqref{ExcitationMatrixEq1} reduce to the usual formulation of the Casida equations in the basis of transitions.
